June 2022 Summary of Project to Determine Asteroid Shape and Density from Flybys

### 2021-09-02 summaries/21-06/21-06.tex
Summaries added back

\section{Introduction}
\subsection{Tidal torque}
The torque $d\bm{\tau}$ imposed on a section of mass $dm$ inside an asteroid by tidal forces is
\begin{equation}
    \frac{d\bm{\tau}}{dm} =\frac{3\mu}{D^3} (\bm{\hat x} yz - \bm{\hat y} xz) = \frac{3\mu}{D^3} \bm{t}(\bm{r})
    \label{eqn:diff-torque}
\end{equation}
where $D$ is the distance between the asteroid, $\mu$ is the gravitational parameter of the planet being orbited. We have oriented our axes such that $\hat {\bm z}$ points towards the planet and $\hat {\bm y}$ points forward in the plane of orbit of the asteroid.

\section{Model}
\subsection{Asteroid model requirements}
\label{sec:requirements}
Any asteroid model with floating paramters designed to describe an asteroid with nonuniform density on flyby with a planet must be well-equipped to perform the following operations

\begin{enumerate}
    \item Rotate the asteroid along its axis of spin
    \item Extract the tidal torque on the asteroid at an instant
    \item Extract the moment of inertia tensor as a function of the parameters of the model
    \item Extract the center of mass of the asteroid as a function of the parameters of the model
\end{enumerate}
Fortunately, the first task requires no extra work; Euler's equations dictate the manner of rotation of a three dimensional body:
\begin{equation}
    I\dot{\bm{\omega}} + \bm \omega \times (I \bm \omega) = \bm \tau
    \label{eqn:euler}
\end{equation}
where $\bm \tau$ is the external torque on the asteroid: tidal torque in this case.

Equation \ref{eqn:diff-torque} is written in the frame where $\hat {\bm z}$ points towards the planet. Assuming the torque $\bm \tau$ in equation \ref{eqn:euler} is also written in this frame, there will be an additional rotation caused by the orbital motion of the asteroid of $\bm \Omega = \bm v \times \hat{\bm z} / D$, where $\bm v$ is the orbital velocity of the asteroid. Note that $\bm \Omega$ is parallel to $\hat x$. From conservation of angular momentum, it follows that $\dot{\bm \Omega} = -2(\bm v \cdot \hat{\bm z})\bm \Omega / D$. The final angular acceleration is then
\begin{equation}
    \dot{\bm \omega} = -\dot{\bm \Omega} + I^{-1}\brackets{\bm \tau - (\bm \omega + \bm \Omega) \times I(\bm \omega + \bm \Omega)}
    \label{eqn:angular-acc}
\end{equation}

>>>>

\section{Introduction}
\subsection{Coordinate frames}
We will make use of three coordinate frames in this project. First is the ``local frame,'' which rotates with the asteroid. We will describe the shape of the asteroid only in the local frame, where it is constant. Next is the ``global frame,'' in which $\hat {\bm z}$ points towards the Earth at all times, $\hat {\bm y}$ points forward in the plane of the orbit, and $\hat {\bm x}$ points perpendicular to the plane of the orbit.

It is important to note that neither of the above two frames are inertial, so we will define a third frame, which is the ``inertial frame.'' We will not discuss the inertial frame much, except as a comparison to derive the phantom forces imposed on the global frame. The inertial frame is defined with $\hat {\bm x}$ pointing perpendicular to the plane of the orbit in the same direction as the $\hat {\bm x}$ of the global frame, so that the inertial frame is just a rotation of the global frame around their common $\hat {\bm x}$

\subsection{Tidal torque}
The torque $d\bm{\tau}$ imposed on a section of mass $dm$ inside an asteroid by tidal forces is
\begin{equation}
    \frac{d\bm{\tau}}{dm} =\frac{3\mu}{D^3} (\bm{\hat x} yz - \bm{\hat y} xz) = \frac{3\mu}{D^3} \bm{t}(\bm{r})
    \label{eqn:diff-torque}
\end{equation}
where $D$ is the distance between the asteroid, $\mu$ is the gravitational parameter of the planet being orbited. This equation is true in the global frame, where $\hat{\bm z}$ points towards the planet.

\section{Model}
\subsection{Asteroid model requirements}
\label{sec:requirements}
Any asteroid model with floating parameters designed to describe an asteroid with nonuniform density on flyby with a planet must be well-equipped to perform the following operations

\begin{enumerate}
    \item Rotate the asteroid along its axis of spin
    \item Extract the tidal torque on the asteroid at an instant
    \item Extract the moment of inertia tensor as a function of the parameters of the model
    \item Extract the center of mass of the asteroid as a function of the parameters of the model
\end{enumerate}
Fortunately, the first task requires no extra work; Euler's equations dictate the manner of rotation of a three dimensional body:
\begin{equation}
    I_\mathcal{I}\dot{\bm{\omega}}_\mathcal{I} + \bm \omega_\mathcal{I} \times (I_\mathcal{I}\bm \omega_\mathcal{I}) = \bm \tau_\mathcal{I}
    \label{eqn:euler-inertial}
\end{equation}
where $\bm \tau_\mathcal{I}$ is the external torque on the asteroid---tidal torque in this case---and the subscript $\mathcal I$ indicates that a quantity is written in the inertial frame. It will be more useful to us to express this equation in the global frame, so we define a rotation matrix $R_x$, which rotates vectors in the global frame to the inertial frame. We also define $\bm \Omega = \bm r \times \bm v/ r^2$ to be the rotation rate of the global frame around the planet, where $\bm r$ is the position and $\bm v$ is the velocity of the asteroid. Since $\bm \Omega$ points along $\hat{\bm x}$, the common axis between the inertial and global frames, we know $R_x \bm\Omega = \bm\Omega$, so that $\bm \omega_\mathcal{I} = R_x(\bm \omega_\mathcal{G} + \Omega)$ where the subscript $\mathcal G$ indicates global coordinates. Similarly, $I_\mathcal{I} = R_x I_\mathcal{G}R_x^T$. Finally, the external torque $\tau$ is simply rotated: $\bm \tau_\mathcal{I} = R_x \bm \tau_\mathcal{G}$.

To simplify equation \ref{eqn:euler-inertial}, we will also need the facts $\dot R_x\bm\omega_\mathcal{G}=R_x(\bm\Omega\times\bm\omega_\mathcal{G})$ and $\dot \Omega = -2(\bm r \cdot \bm v / r^2)\bm \Omega$. Taking this into account, we have
\begin{equation}
    \dot{\bm{\omega}} = I^{-1}\brackets{\bm \tau - (\bm \omega + \bm \Omega) \times I (\bm \omega + \bm \Omega)} + 2 \bm \Omega \frac{\bm r \cdot \bm v}{r^2} - \bm \Omega \times \bm \omega
\end{equation}
where all the quantities are now in global coordinates.
